---
title: "Assignment"
author: "DataBrew"
output: 
    html_document:
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('fivethirtyeight')
library(fivethirtyeight)
library(tidyverse)
```


## 538 data (drug use)


```{r}
# create an object called 'du' and assign to it the 'drug_use' dataset
du <- drug_use
```

## plotting exercises

1. Create a histogram of one variable

2. Create a density plot of one variable.

3. Create a histogram of with two or more variables.

4. Create a barplot using two variables. 

5. Create a scatter plot using two variables.

6. Create a scatter plot using three variables 

7. Create a scatter plot using four variables.

## creating new variables from our data

1. Create a variable call age_category. The categories will be "old" and 
"young". 

2. What is the average marijuana use by age_category

3. 

## spatial data

```{r}
# read in world cities data 
world_cities <- read_csv('data/simplemaps-worldcities-basic.csv')

```

1. Create a map of all the cities in the world.

2. Create a map of all the cities in Mozambique

3. Create a map of all the cities in Mozambique and show population

4. Create a map of all the cities in Kenya, France, and China and show population

5. 