---
title: "Exercises"
author: "DataBrew"
output: 
    html_document:
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Thursday {.tabset .tabset-fade}


```{r, include = FALSE}
library(tidyverse)
```

### Getting data

1. Load up some libraries

```{r, include = FALSE, echo = FALSE}
library(raster)
```

```{r, eval = FALSE, echo = TRUE}
library(raster)
```

2. Find the 3 letter ISO code of a country you like.s Here is a list of codes: https://www.iban.com/country-codes

3. Create an object called `iso`. This will be 3 letter ISO code. For example, here is Tanzania (but you can do a different country)

```{r}
iso <- 'TZA'
```

4. Get the data for your country into an object called `x`. Set the level to `0` if you want for only the country, set it higher if you want more detailed

```{r}
x <- getData(country = iso, level = 2)
```

5. Make a very simple plot, just to confirm that it all worked.

```{r}
plot(x)
```



### Spatial 0

1. Load up some packages

```{r, include = FALSE, echo = FALSE}
library(choroplethr)
library(choroplethrMaps)
library(tidyverse)
library(gapminder)
```

```{r, eval = FALSE, echo = TRUE}
library(choroplethr)
library(choroplethrMaps)
library(tidyverse)
library(gapminder)
```



2. Get the gapminder data. Name it `gm`

```{r}
gm <- gapminder
```

3. Rename some things

```{r}
gm <- gm %>%
  rename(region = country,
         value = lifeExp) %>%
  mutate(region = tolower(region))
```

4. Filter for only 2007

```{r}
gm_choro <- gm %>% filter(year == 2007)
```

5. Plot

```{r, message=FALSE, eval = FALSE}
country_choropleth(gm_choro)
```

6. Make only 4 colors

```{r, eval = FALSE}
country_choropleth(gm_choro, num_colors = 4)
```

7. Extra/advanced. Clean up the mismatched country names and make the charts prettier.



### Spatial 1

1. Load up some useful packages

```{r, message = FALSE}
library(tidyverse)
library(rgdal)
library(raster)
library(sp)
```

2. Get some Mozambican geographic data from the internet and assign it to an object named `moz0`.

```{r, message = FALSE}
library(raster)
library(leaflet)
moz0 <- getData(country = 'MOZ', level = 0)
```

3. Create a basic leaflet plot.

```{r}
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = moz0)
```

4. Get some more Mozambican data. This time, at the province level (ie, level 1).

```{r, message = FALSE}
moz1 <- getData(country = 'MOZ', level = 1)
```


5. Create a leaflet map of the moz1 object (see question 3 if you need help).

```{r, eval = FALSE}
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = moz1)
```


6. Filter to only keep Gaza

```{r}
gaza <- moz1[moz1$NAME_1 == 'Gaza',]
```

7. Make a leaflet map of it.

```{r, eval = FALSE}
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = gaza)
```

8. Read in some data on health posts.


```{r, eval = FALSE, echo = TRUE}
us <- read_csv("https://raw.githubusercontent.com/databrew/databrew.github.io/master/us.csv")
```


```{r, echo = FALSE}
us <- read_csv("https://raw.githubusercontent.com/databrew/databrew.github.io/master/us.csv", col_types = cols())
```

9. Subset the health posts to just include those in Gaza

```{r}
gaza_us <- us %>% filter(province == 'GAZA')
```

10. Make a plot with both the polygon of the Gaza province and the health posts of the Gaza province

```{r, eval = FALSE, echo = FALSE}
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = gaza) %>%
  addCircleMarkers(data = gaza_us, col = 'yellow')
```


### Spatial 2

```{r, eval = FALSE, echo = FALSE}
library(bohemia)
library(cism)
mopeia_us <- us %>% filter(district == 'MOPEIA')
save(mop2, mop3, mopeia_roads, mopeia_water, mopeia_waterways,
     mopeia2, mopeia3,
     man3, man2,
     mopeia_us,
     file = 'mop.RData')
```

1. Load up some spatial data from the internet by running the following:

```{r}
load(url('https://github.com/databrew/maputo/raw/master/day3/mop.RData'))
```

2. Make a plot of the district borders (use the mop2 object)

```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon()
```


3. Make the border of the district black, and the inside of the district grey.

```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey')
```

4. Add `theme_map()` to remove the background

```{r, echo = FALSE}
library(ggthemes)
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey') +
  theme_map()
```

5. Add a title, subtitle, and caption


```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey') +
    theme_map() +
  labs(title = 'The district of Mopeia',
       subtitle = 'Map',
       caption = 'Produced by Carla')
```

6. Add rivers by adding `geom_path` with the `mopeia_waterways` object. Color the water blue.


```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey') +
    theme_map() +
  labs(title = 'The district of Mopeia',
       subtitle = 'Map',
       caption = 'Produced by Carla') +
  geom_path(data = mopeia_waterways,
            color = 'blue')
```



7. Add roads using the `mopeia_roads` object. Color them brown


```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey') +
    theme_map() +
  labs(title = 'The district of Mopeia',
       subtitle = 'Map',
       caption = 'Produced by Carla') +
  geom_path(data = mopeia_waterways,
            color = 'blue') +
  geom_path(data = mopeia_roads,
            color = 'brown')
```

8. Add points for the unidades sanitarias. Color them yellow. You'll need to use the `mopeia_us` data and add `group=NA` to within your call to `geom_point()`.

```{r, echo = FALSE}
ggplot(data = mop2,
       aes(x = long,
           y = lat,
           group = group)) +
  geom_polygon(color = 'black',
               fill = 'grey') +
    theme_map() +
  labs(title = 'The district of Mopeia',
       subtitle = 'Map',
       caption = 'Produced by Carla') +
  geom_path(data = mopeia_waterways,
            color = 'blue') +
  geom_path(data = mopeia_roads,
            color = 'brown') +
  geom_point(data = mopeia_us,
             aes(x = longitude,
                 y = latitude),
             group = NA,
             color = 'yellow')
```

9. ADVANCED: Reproduce the above map, but in an interactive version using leaflet. Some hints:

- You'll need the `addPolylines` function for roads and water.
- `weight` makes things thinner
- `opacity` controls alpha

```{r, echo = FALSE}
leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addPolygons(data = mopeia2) %>%
  addPolylines(data = mopeia_roads,
                           color = 'brown',
               weight = 2,
               opacity = 1)%>%
  addPolylines(data = mopeia_waterways,
                           color = 'blue',
               weight = 1,
               opacity = 1) %>%
  addCircleMarkers(data = mopeia_us,
                   color = 'yellow')
```


### Spatial 3


1. Load up some packages

```{r}
library(tidyverse)
```

2. Create an object called `terra` like this:

```{r, eval = FALSE, echo = FALSE}
terra <- map_data("world")
change <- function(ff, tt){
  terra$country[terra$country == ff] <- tt
}
terra$country <- as.character(terra$region)
terra$country[terra$country == 'Democratic Republic of the Congo'] <- "Congo, Dem. Rep."
terra$country[terra$country == 'Republic of Congo'] <- "Congo, Rep."
terra$country[terra$country == "Ivory Coast"] <- "Cote d'Ivore"
terra$country[terra$country == "North Korea"] <- "Korea, Dem. Rep."
terra$country[terra$country == "South Korea"] <- "Korea, Rep."
terra$country[terra$country == "Slovakia"] <- "Slovak Republic"
terra$country[terra$country == "Trinidad"] <- "Trinidad and Tobago"
terra$country[terra$country == "UK"] <- "United Kingdom"
terra$country[terra$country == "USA"] <- "United States"
terra$country[terra$country == "Palestine"] <- "West Bank and Gaza"
terra$country[terra$country == "Yemen"] <- "Yemen, Rep."
write_csv(terra, 'terra.csv')
```

```{r, include = FALSE}
terra <- read_csv('https://github.com/databrew/maputo/blob/master/day4/terra.csv?raw=true')
```

```{r, echo = TRUE, eval = FALSE}
terra <- read_csv('https://github.com/databrew/maputo/blob/master/day4/terra.csv?raw=true')
```

3. Plot `terra`

4. Plot `terra` with clear borders between each country

5. Create an object named `mozy`. This will just be Mozambique.

6. Plot `mozy`.

7. Plot `mozy` in an interactive map.

8. Load up the gapminder data as an object called `gm`.

```{r}
gm <- gapminder::gapminder
```

9. Filter the gapminder data to only include data from 2007. Call this `recent`.


10. Perform a join between `terra` and `recent` so as to associate geographic data with health/wealth data. Call the new objected `combinado`

```{r, echo = FALSE, eval = FALSE}
combinado <- left_join(terra, recent)
```

11. Plot `combinado` using ggplot, but filling each country as a function of gdpPerCap

```{r, echo = FALSE, eval = FALSE}
ggplot(data = combinado,
       aes(x = long,
           y = lat,
           group = group,
           fill = gdpPercap)) +
  geom_polygon() +
  scale_fill_gradientn(colors = c('red', 'white', 'blue'))
```

12. Let's alter the color scale by adding the following:

```{r, echo = TRUE, eval = FALSE}
scale_fill_gradientn(colors = c('red', 'white', 'blue'))
```

13. Change the colors, change the theme, add borders, etc.

### Spatial 4

1. Load up some packages

```{r}
library(leaflet)
library(sp)
library(tidyverse)
library(raster)
```

2. Load the violence data from yeterday

```{r, include = FALSE}
violence <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/moz.csv?raw=true")
```


```{r, echo = TRUE, eval = FALSE}
violence <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/moz.csv?raw=true")
```


3. Load a shapefile of Mozambique

```{r}
moz1 <- getData(country = 'MOZ', level = 1)
```

4. In `violence`, create a new column named `province`. This will be simply the `admin1` column copied.

5. In `moz1`, create a new column called `province`. This will be simply the `NAME_1` column copied.

6. In `moz1`, change the word "Nassa" to "Niassa" in the `province` column.

7. Create an object called `bp`. This will be the number of fatalities in violent events by province, only since 2010.

```{r, echo = FALSE, eval = FALSE}
bp <- violence %>% filter(year >= 2010) %>% group_by(province = admin1) %>% summarise(deaths = sum(fatalities))
moz1$province <- moz1$NAME_1
moz1$province <- ifelse(moz1$province == 'Nassa', 'Niassa', moz1$province)
```

8. Join `moz1` and `bp`. **NOTE**, since `moz1` is not a normal dataframe (it's a "SpatialPolygonsDataFrame"), we have to specifically use it's "data" component. The below is how to do this.

```{r, echo = TRUE, eval = FALSE}
moz1@data <- left_join(x = moz1@data,
                       y = bp,
                       by = 'province')
```

9. Make an interactive map with moz1

```{r, echo = FALSE, eval = FALSE}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = moz1)
```

10. Create a choropleth of the deaths data

```{r, echo = TRUE, eval = FALSE}
pal <- colorNumeric("Spectral", domain = range(moz1@data$deaths))

leaflet() %>%
  addTiles() %>%
  addPolygons(data = moz1,
              fillColor = ~pal(deaths),
              weight = 1,
              color = 'black',
              opacity = 0.9)
```

11. Add a legend with the following line:

```{r, echo = TRUE, eval = FALSE}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = moz1,
              fillColor = ~pal(deaths),
              weight = 1,
              color = 'black',
              opacity = 0.9) %>%
  addLegend(pal=pal, values=moz1$deaths, opacity=0.9, title = "Deaths since 2010", position = "bottomright" )

```



### Weather

1. Read in some Mozambican weather data by running the following

```{r, include = FALSE, echo = FALSE}
library(tidyverse)
weather <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/mozambican_weather.csv?raw=true")
```

```{r, eval = FALSE, echo = TRUE}
library(tidyverse)
weather <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/mozambican_weather.csv?raw=true")
```

2. Take a look at the top of the data. 

3. Create an object called `x`. This will be the `weather` data, then summarized to tell us the minimum and maximum dates.

4. Create an object called `y`. This will be the `weather` data, then summarized to tell us the maximum `temp_max` and the mimimum `temp_min`.

5. Create an object called `z`. This will be the `weather` data, but filtered to only include the year 2014 and only the district of "MATOLA".

6. Make a plot of `z` with the `date` on the x-axis and the average temperature (`temp`) on the y-axis.

7. Make a plot of `z` showing the distribution of precipitation.

8. Make a plot showing the distribution of the maximum temperature, but faceted by district.

```{r, eval = FALSE, echo = FALSE}
ggplot(data = weather,
       aes(x = temp_max)) +
  geom_density() +
  facet_wrap(~district)
```

9. Color the lines in the above red.

10. Add another distribution the above chart with the minimum temperature. Color it blue. To do this you will need to add another `geom_density` line to your chart, and you can include another `aes()` section within that `geom_density` function.


### Making long data

1. Read in some Mozambican weather data by running the following

```{r, include = FALSE, echo = FALSE}
library(tidyverse)
weather <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/mozambican_weather.csv?raw=true")
```

```{r, eval = FALSE, echo = TRUE}
library(tidyverse)
weather <- read_csv("https://github.com/databrew/maputo/blob/master/day3/data/mozambican_weather.csv?raw=true")
```

2. Take a look at the top of the data. 

3. Create an object called `pd`. This will be some data we want to plot. It will include only the columns we want to plot. For example.

```{r}
pd <- weather %>%
  dplyr::select(district, date, temp_max, temp_min)
```

4. Make the data "long" using `gather`

```{r}
pdl <- pd %>%
  gather(key, value, temp_max:temp_min)
```

5. Make a plot of temperature with the color showing the `key` and paneling/faceting for each district.

6. Read in the gapminder data

```{r}
library(gapminder)
gm <- gapminder
```

7. Create an object called `x`. This will only be data for countries in Oceania

```{r, echo = FALSE, eval = TRUE}
x <- gm %>% filter(continent == 'Oceania')
```

8. Make the data "long" by gathering the lifeExp, pop, and gdpPercap variables.

```{r}
x <- x %>% gather(key, value, lifeExp:gdpPercap)
```

9. Plot the `x` data with year on the x-axis, the `value` on the y-axis, colored by country, and panelled/faceted by `key`. Add `scales = 'free_y'` to within the `facet_wrap` function in order to have "loose" scales.

10. Filter `x` to remove population. Plot again (using the same code from 9)

### Switzerland

1. Load up data on health and sociodemographic factors from Switzerland.

```{r}
library(tidyverse)
data('swiss')
swiss <- swiss
```

2. Look at the `cabeça` of the data. Describe it.

3. Make a chart showing the association between Agriculture rate and Fertility rate. This should be a scatterplot. You'll use `geom_point`.

```{r, eval = FALSE, echo = FALSE}
ggplot(data = swiss,
       aes(x = Agriculture,
           y = Fertility)) +
  geom_point()
```

4. Describe the association.

5. Add the argument `geom_smooth()` to the above chart. What happens?

```{r, eval = FALSE, echo = FALSE}
ggplot(data = swiss,
       aes(x = Agriculture,
           y = Fertility)) +
  geom_point() +
  geom_smooth()
```


6. Run the following code to create a variable called `canton` from the `row.names` of the data:

```{r}
swiss$canton <- row.names(swiss)
```

7. Create an object called `x`. This will be the `swiss` data, but only keeping those cantons which are more than 80% catholic.

```{r, eval = FALSE, echo = FALSE}
x <- swiss %>% filter(Catholic > 80)
```

8. How many rows are in `x`?

9. Go back to your `swiss` data. Is there an association between `Education` and `Catholicism`? Show it in a chart.

10. Create a variable a new variable in `swiss` called `edu`, which indicates whether the canton is highly-educated (at or above the average) or has low education (below the average). Here's the code to do this:

```{r}
swiss <- swiss %>%
  mutate(edu = ifelse(Education >= mean(Education),
                      "High education",
                      "Low education"))
```

11. Create a variable in `swiss` called `fert`. This variable should show whether the canton has high fertility or low fertility. Use the code in the previous question as a guide.

```{r, eval = TRUE, echo = FALSE}
swiss <- swiss %>%
  mutate(fert = ifelse(Fertility >= mean(Fertility),
                      "High fertility",
                      "Low fertility"))
```

12. Create an object called `y`. This will be the `swiss` data, grouped by both your new `edu` and `fert` variables, and then summarized to tell the average infant mortality rate for each subgroup.

```{r, echo = FALSE}
y <- swiss %>%
  group_by(fert, edu) %>%
  summarise(avg = mean(Infant.Mortality))
```

13. Make a barplot of the data in y. Add the argument `position = 'dodge` to within `geom_bar()` to make the bars not stack

```{r, echo = FALSE, eval = FALSE}
ggplot(data = y,
       aes(x = fert, y= avg, fill = edu)) +
  geom_bar(stat = 'identity', position = 'dodge')
```

14. Interpret the above chart. Which group has the highest infant mortality? Which one has the lowest?


15. Arrange the data by infant mortality. Which canton has the lowest?

16. What is the maximum infant mortality for all Swiss canons?

17. What is the maximum infant mortality for canons with high education?

18. What is the average infant mortality for canons with high fertility.

19. Make a plot showing the relationship between fertility (on the x-axis) and infant mortality (on the y-axis).

```{r, echo = FALSE, eval = FALSE}
ggplot(data = swiss,
       aes(x = Fertility,
           y = Infant.Mortality)) +
  geom_point()
```

20. Add `color = edu` to within the `aes()` section of the above chart.

```{r, echo = FALSE, eval = FALSE}
ggplot(data = swiss,
       aes(x = Fertility,
           y = Infant.Mortality,
           color = edu)) +
  geom_point()
```


21. Make a "density" plot of fertility.

22. Create a new variable in `swiss` using `mutate`. The new variable will be named `ag`. This should be whether a canton is very agricultural or not (ie, more than 50%). To do this:

```{r}
swiss <- swiss %>%
  mutate(ag = ifelse(Agriculture > 50, 'Very agricultural',
                     'Not very agricultural'))
```

23. Create an overlapping density chart of infant mortality in which the `fill` shows whether a place is very agricultural or not.

24. Create a new variable in your data called `cath`. This should whether or not a canton is very Catholic (ie, > 50%).

```{r, eval = FALSE, echo = FALSE}
swiss <- swiss %>%
  mutate(cath = ifelse(Catholic > 50, 'Very Catholic', 'Not very Catholic'))
```

25. Create an overlapping density chart of fertility in which the `fill` is whether a place is very Catholic or not.


26. Use `ggridges` to create a ridgeline chart of the distribution of infant mortality by your `cath` variable (here's the code to help)

```{r, eval = FALSE, echo = TRUE}
library(ggridges)
ggplot(data = swiss,
       aes(x = Infant.Mortality,
           y = cath)) +
  ggridges::geom_density_ridges()
```

27. Create a ridgeline chart (like above) showing fertility instead of infant mortality.

```{r, eval = FALSE, echo = FALSE}
library(ggridges)
ggplot(data = swiss,
       aes(x = Fertility,
           y = cath)) +
  ggridges::geom_density_ridges()
```

28. Which areas have more children: Catholic or non-Catholic areas?

29. Use the `cut` function to generate a 4-category categorical variable of the `Education` variable. Call this new variable `educat`. Note that `cut` takes two arguments: first, the variable you are cutting; second, the number of categories you want.

```{r, eval = TRUE, echo = FALSE}
swiss <- swiss %>%
  mutate(educat = cut(Education, 4))
```

30. Make a ridgeline plot of infant mortality's distribution by your new `educat` variable.

```{r, eval = FALSE, echo = FALSE}
ggplot(data = swiss,
       aes(x = Infant.Mortality,
           y = educat)) +
  geom_density_ridges(alpha = 0.6)
```

31. Can you create the following plot? You'll need to add `facet_wrap(~educat)`...

```{r}
ggplot(data = swiss,
       aes(x = Catholic,
           y = Infant.Mortality,
           size = Fertility)) +
  facet_wrap(~educat) +
  geom_point(alpha = 0.6)
```

32. What percentage of high-Catholicism cantons are agricultural?

33. What percentage of low fertility canons are Catholic?



### Time data

1. Load up some data about the weight of chickens over time. Create an object called `frangos`.

```{r}
data('ChickWeight')
frangos <- ChickWeight
```

2 How many columns does the data have?

3 How many rows does the data have?

4 What are the variable types (quantitative/numeric or categorical)?

5 Create a point chart showing weight on the x-axis and time on the y-axis.

6 Create an object called `frango1`. This should be just the data for chicken number 1 (ie, 1 in the `Chick` column).

7 Chart the weight of chicken 1 over time using `geom_point()`

8 Chart the weight of chicken 1 over time using `geom_line()`

9 Chart the weight of chicken 1 over time using `geom_area()`

10 Chart the weight of _all_ chickens over time using `geom_line()`. Make the color of each line different for each chicken.

11 Create an object called `zero`. This should be the frangos only at day 0 (ie, when they are born).

12 Make a histogram of chicken's weights at day 0.

13 Make a density plot of chicken's weights at day 0.

14 Make a violin plot of chicken's weights at day 0. On the x-axis, put the `Diet` type. On the y-axis, put the weight.

15 Add points to the violin plot

### correlation plots

- between -1 and 1
- positive: as x increases, y increases
- negative: as x increases, y decreases
- 0: no correlation

### scatter plot 

```{r}
library(gapminder)
library(tidyverse)

dat <- gapminder

ggplot(dat, aes(gdpPercap, lifeExp)) + geom_point()


```

### add correlation to plot


```{r}
corr <- cor(dat$gdpPercap, dat$lifeExp)
ggplot(dat, aes(gdpPercap, lifeExp)) + geom_point() +
  labs(title = corr)


```


```{r}
corr <- round(corr, digits = 2)
ggplot(dat, aes(gdpPercap, lifeExp)) + geom_point() +
  labs(title = corr)


```

### correlation matrix

```{r}
install.packages("corrplot")
library(corrplot)
cars <- mtcars
?mtcars
# create correlation matrix
m <-cor(mtcars)
head(round(m,2))

```

### plot the correlation matrix

```{r}
corrplot(m, method="circle")

```

```{r}
corrplot(m, method="pie")

```

```{r}
corrplot(m, method="color")

```

```{r}
corrplot(m, method="number")

```

```{r}
corrplot(m, method="upper")

```

```{r}
corrplot(m, method="lower")

```


## Time series 

```{r}
library(tidyverse)
dat <-economics
ggplot(data = dat, aes(x = date, y = psavert)) +
  geom_line() +
  labs(title = "Personal Savings Rate",
       x = "Date",
       y = "Personal Savings Rate")

```

## using geom_smooth() to fit a linear model over the data
```{r}
ggplot(dat, aes(x = date, y = psavert)) +
  geom_line(color = "indianred3", 
            size=1 ) +
  geom_smooth(method = 'lm') +
  labs(title = "Personal Savings Rate",
       subtitle = "1967 to 2015",
       x = "",
       y = "Personal Savings Rate") +
  theme_minimal()
```

## Loess - Locally weighted least squares regression
```{r}
ggplot(dat, aes(x = date, y = psavert)) +
  geom_line(color = "indianred3", 
            size=1 ) +
  geom_smooth(method = 'loess') +
  labs(title = "Personal Savings Rate",
       subtitle = "1967 to 2015",
       x = "",
       y = "Personal Savings Rate") +
  theme_minimal()
```

## gather data into long form
```{r}
long_dat <- gather(data = dat, variable, value, -date) %>% 
 filter(variable == 'pce' | variable == 'unemploy')

head(long_dat)

```

## plot all variables
```{r}
ggplot(data = long_dat, aes(x = date, y = value, color = variable)) +
  geom_line(size=1) +
  labs(title = "Personal Savings Rate",
       subtitle = "1967 to 2015",
       x = "",
       y = "Personal Savings Rate") +
  theme_minimal()

```

## Area charts
```{r}
# basic area chart
ggplot(dat, aes(x = date, y = psavert)) +
  geom_area(fill="lightblue", color="black") +
  labs(title = "Personal Savings Rate",
       x = "Date",
       y = "Personal Savings Rate")
```


## Dumbell chart
```{r}
# subset data
# install.packages('ggalt')
# install.packages('devtools')
devtools::install_github("hrbrmstr/ggalt", ref = "noproj")

library(ggalt)
library(gapminder)
dat <- gapminder %>% filter(continent == "Americas" &
                                      year == 1952 | year == 2007) %>%
  select(country, year, lifeExp)

# convert data to wide format
dat_wide <- spread(dat, year, lifeExp)
names(dat_wide) <- c("country", "y1952", "y2007")

# create dumbbell plot
ggplot(dat_wide, aes(y = country,
                          x = y1952,
                          xend = y2007)) +  
  geom_dumbbell()

```

## add titles and labels
```{r}
ggplot(dat_wide, 
       aes(y = reorder(country, y1952),
           x = y1952,
           xend = y2007)) +  
  geom_dumbbell(size = 1.2,
                size_x = 3, 
                size_xend = 3,
                colour = "grey", 
                colour_x = "blue", 
                colour_xend = "red") +
  theme_minimal() + 
  labs(title = "Change in Life Expectancy",
       subtitle = "1952 to 2007",
       x = "Life Expectancy (years)",
       y = "")
```

## slope charts

```{r}
library(CGPfunctions)
install.packages('CGPfunctions')
# Select Central American countries data 
# for 1992, 1997, 2002, and 2007

df <- gapminder %>%
  filter(year %in% c(1992, 1997, 2002, 2007) &
           country %in% c("Panama", "Costa Rica", 
                          "Nicaragua", "Honduras", 
                          "El Salvador", "Guatemala",
                          "Belize")) %>%
  mutate(year = factor(year),
         lifeExp = round(lifeExp)) 

# create slope graph
newggslopegraph(df, year, lifeExp, country) +
  labs(title="Life Expectancy by Country", 
       subtitle="Central America", 
       caption="source: gapminder")
```
=======
### Markdown

1. Install these packages if you don't yet have them

```{}
knitr
rmarkdown
DT
```

2. In the upper left of RStudio, click "File", then "New File", then "R Markdown".

3. Name your file: "Mapping with R"

4. In your new file, run CTRL + s in order to save. Save it as "mapping" (for example).

5. Delete lines 11 until the end.

6. Write the following in order to create a "header" section:

```

## How to get data

```

7. Move your cursor below that section, and then click on "Insert" at the top of the code window. From there, click on "R".

8. You now have a section for writing R code. In that section, write:

```{r}
library(raster)
library(leaflet)
library(tidyverse)
```

9. Click "Knit" in the top of the window in order to compile the document.

10. Now add more text and R sections to create a complete guide for mapping in R.
