---
title: "Data Visualization using R"
subtitle: "Day 3"
author: "Arsénio Nhacolo, Joe Brew and Ben Brew"
date: "02-06/12/2019"
output: 
  ioslides_presentation:
    logo: www/logo_full.png
    css: www/cism_styles.css
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
npac <- c("ggplot2","plotly" , "dplyr", "tidyr", 
        "gapminder", "ggmap", 
        "knitr", "kableExtra")
for (pk in npac){
  library(pk,character.only = TRUE)
}

```




# Day 3

## Today's agenda

### Morning
- Ben: **Multivariate graphs** - Introduction
- Joe: **Multivariate graphs** - Preparing the data (dplyr, grouping, etc.).
- Arsenio: **Multivariate graphs** - Practical instruction
- All: **Multivariate graphs** - Exercises

### Afternoon
- Joe: **Mapping** - Introduction  
- Arsenio: **Mapping** - Point maps / dot density maps
- Ben: **Mapping** - Choropleth maps  
- All: **Mapping** - Exercises


# Multivariate graphs

## Multivariate graphs|Introduction (Ben)

- How do we show 3 (or more) variables on a 2 dimensional plot?

## two variable bar plot
```{r}
library(tidyverse)
library(ggthemes)

# for the afternoon, we need to install rgdal.
# install.packages('leaflet')
dat <- read.csv('../../police_shootings/data/fatal-police-shootings-data.csv')

new_dat <- dat %>% group_by(flee) %>% summarise(counts = n()) %>% filter(flee != '')
ggplot(data = new_dat, aes(x = flee, counts)) + 
  geom_bar(stat = 'identity') + labs(x = '', y = 'Counts', title = 'People killed by police') + theme_clean()

```

## Other variables to

```{r}
summary(dat)

```

## Add 3rd variable
```{r}

new_dat <- dat %>% group_by(flee, threat_level) %>% summarise(counts = n()) %>% filter(flee != '')

ggplot(data = new_dat, aes(x = flee, y = counts, fill = threat_level)) + 
  geom_bar(stat = 'identity', position = 'dodge') + labs(x = '', y = 'Counts', title = 'People killed by police') + theme_clean()

```

## Numeric data (2 variables)

- Life expectancy and gdp per capita in 2007


```{r}
dat <- gapminder
dat <- dat %>% filter(year == '2007')
ggplot(data = dat, aes(gdpPercap, lifeExp)) + 
  geom_point(size = 2, alpha = 0.4) +
  labs(x = 'GDP per capita', 
       y = 'Life expectancy',
       title = 'GDP per capita and Life expectancy') + theme_clean()

```

## Numeric data (3 variables)

- Life expectancy and gdp per capita in 2007
- The size of the points represents population

```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% filter(year == '2007')
ggplot(data = dat, aes(gdpPercap, lifeExp)) + 
  geom_point(aes(size = pop), alpha = 0.4) +
  labs(x = 'GDP per capita', 
       y = 'Life expectancy',
       title = 'GDP per capita and Life expectancy') + theme_clean()
```

## Numeric data (4 variables)

- Life expectancy and gdp per capita in 2007
- The size of the points represents population

```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% filter(year == '2007')
ggplot(data = dat, aes(gdpPercap, lifeExp)) + 
  geom_point(aes(size = pop, color = continent), alpha = 0.4) +
  labs(x = 'GDP per capita', 
       y = 'Life expectancy',
       title = 'GDP per capita and Life expectancy') + theme_clean()
```

## Another example


```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% group_by(year, continent) %>% summarise(avg_gdp = mean(gdpPercap),
                                                       avg_life_exp = mean(lifeExp),
                                                       avg_pop = mean(pop))
# dat <- dat %>% filter(continent == 'Oceania')
ggplot(data = dat, aes(year, avg_pop)) +
  geom_point(aes(color = continent)) +
  geom_line(aes(color = continent)) +
  labs(x = 'Year', 
       y = 'Population',
       title = 'Year and population by continent') + theme_clean()
```

## Another example

```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% group_by(year, continent) %>% summarise(avg_gdp = mean(gdpPercap),
                                                       avg_life_exp = mean(lifeExp),
                                                       avg_pop = mean(pop))
# dat <- dat %>% filter(continent == 'Oceania')
ggplot(data = dat, aes(year, avg_pop)) +
  geom_point(aes(color = continent, size = avg_gdp)) +
  geom_line(aes(color = continent)) +
  labs(x = 'Year', 
       y = 'Population',
       title = 'Year and population by continent') + theme_clean()
```

## Another example

```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% group_by(year, continent) %>% summarise(avg_gdp = mean(gdpPercap),
                                                       avg_life_exp = mean(lifeExp),
                                                       avg_pop = mean(pop))
# dat <- dat %>% filter(continent == 'Oceania')
ggplot(data = dat, aes(year, avg_pop)) +
  geom_point(aes(color = continent, size = avg_life_exp)) +
  geom_line(aes(color = continent)) +
  labs(x = 'Year', 
       y = 'Population',
       title = 'Year and population by continent') + theme_clean()
```

## Asia example
```{r}
options(scipen = '999')
dat <- gapminder
dat <- dat %>% group_by(year, country) %>% filter(continent == 'Asia' & country %in% c('China', 'Japan', 'Singapore', 'Korea, Rep.')) %>% summarise(gdp = mean(gdpPercap),
                                                       life_exp = mean(lifeExp),
                                                       pop = mean(pop))
# dat <- dat %>% filter(continent == 'Oceania')
ggplot(data = dat, aes(year, gdp)) +
  geom_point(aes(color = country, size = pop)) +
  geom_line(aes(color = country)) +
  labs(x = 'Year', 
       y = 'GDP per capita',
       title = 'Year and GDP') + theme_clean()

```

## Multivariate graphs|Data preparation (Joe)

## Multivariate graphs|Practical instruction (Arsenio)

## Multivariate graphs

* Display the relationships among three or more variables in a single graph
* Two common methods for accommodating multiple variables: grouping and faceting

## Multivariate graphs|Grouping

* The values of the first two variables are mapped to the x and y axes
* Additional variables are mapped to other visual characteristics such as color, shape, size, line type, and transparency 

## Multivariate graphs|Grouping -- 3 variables (colour)

```{r, eval=FALSE}
# Plot experience vs. salary (color represents rank)
ggplot(salaries, aes(x = yrs_since_phd, 
                     y = salary, 
                     color=rank)) +
  geom_point() +
  labs(title = "Academic salary by rank and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank") +
  theme_pubr()
```

## Multivariate graphs|Grouping -- 3 variables (colour)

```{r, echo=FALSE, fig.width=9, fig.height=4.3}
salaries <- rio::import("data/salaries.rds")
# Plot experience vs. salary (color represents rank)
ggplot(salaries, aes(x = yrs_since_phd, 
                     y = salary, 
                     color=rank)) +
  geom_point() +
  labs(title = "Academic salary by rank and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank") +
  theme_pubr()
```

## Multivariate graphs|Grouping -- 4 variables (colour and shape)

```{r, eval=FALSE}
# Plot experience vs. salary (color represents rank, shape represents sex)
ggplot(salaries, 
       aes(x = yrs_since_phd, 
           y = salary, 
           color = rank, 
           shape = sex)) +
  geom_point(size = 3, 
             alpha = .6) +
  labs(title = "Academic salary by rank, sex, and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank",
       shape = "Gender") +
  theme_pubr()
```

## Multivariate graphs|Grouping -- 4 variables (colour and shape)

```{r, echo=FALSE, fig.width=9, fig.height=4.3}
# Plot experience vs. salary (color represents rank, shape represents sex)
ggplot(salaries, 
       aes(x = yrs_since_phd, 
           y = salary, 
           color = rank, 
           shape = sex)) +
  geom_point(size = 3, 
             alpha = .6) +
  labs(title = "Academic salary by rank, sex, and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank",
       shape = "Gender") +
  theme_pubr()
```


## Multivariate graphs|Grouping -- 4 variables (colour and size)

```{r, eval=FALSE}
# Plot experience vs. salary (color represents rank and size represents service)
ggplot(salaries, 
       aes(x = yrs_since_phd, 
           y = salary, 
           color = rank, 
           size = yrs_service)) +
  geom_point(size = 3, 
             alpha = .6) +
  labs(title = "Academic salary by rank, years of service, and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank",
       size = "Years of service") +
  theme_pubr()
```

## Multivariate graphs|Grouping -- 4 variables (colour and size)

```{r, echo=FALSE, fig.width=9, fig.height=4.3}
# Plot experience vs. salary (color represents rank and size represents service)
ggplot(salaries, 
       aes(x = yrs_since_phd, 
           y = salary, 
           color = rank, 
           size = yrs_service)) +
  geom_point(alpha = .6) +
  labs(title = "Academic salary by rank, years of service, and years since degree",
       x = "Years since PhD",
       y = "Salary ($)",
       color = "Rank",
       size = "Years of service") +
  theme_pubr()
```

## Multivariate graphs|Faceting

* In faceting, a graph consists of several separate plots or small multiples, one for each level of a third variable, or combination of variables

```{r, eval=FALSE}
# Plot salary histograms by rank
ggplot(salaries, aes(x = salary)) +
  geom_histogram(fill = "lightseagreen",
                 color = "white") +
  facet_wrap(~rank, ncol = 1) +
  labs(title = "Salary histograms by rank") +
  theme_pubr()
```

## Multivariate graphs|Faceting -- 3 variables

```{r, echo=FALSE, fig.width=9, fig.height=4.3, message=FALSE}
# Plot salary histograms by rank
ggplot(salaries, aes(x = salary)) +
  geom_histogram(fill = "lightseagreen",
                 color = "white") +
  facet_wrap(~rank, ncol = 1) +
  labs(title = "Salary histograms by rank") +
  theme_pubr()
```

## Multivariate graphs|Faceting -- 4 variables

```{r, eval=FALSE}
# Plot salary histograms by rank and sex
ggplot(salaries, aes(x = salary/1000)) +
  geom_histogram(color = "white",
                 fill = "lightseagreen") +
  facet_grid(sex ~ rank) +
  labs(title = "Salary histograms by sex and rank",
       x = "Salary ($1000)") +
  theme_pubr()
```

## Multivariate graphs|Faceting -- 4 variables

```{r, echo=FALSE, fig.width=9, fig.height=4.3, message=FALSE}
# Plot salary histograms by rank and sex
ggplot(salaries, aes(x = salary/1000)) +
  geom_histogram(color = "white",
                 fill = "lightseagreen") +
  facet_grid(sex ~ rank) +
  labs(title = "Salary histograms by sex and rank",
       x = "Salary ($1000)") +
  theme_pubr()
```

## Multivariate graphs|Combining grouping and faceting

* E.g.: use Mean/SE plots and faceting to compare the salaries of male and female professors, within rank and discipline

```{r}
# Calculate means and standard erroes by sex, rank and discipline
library(dplyr)
sal_srd <- salaries %>%
  group_by(sex, rank, discipline) %>%
  summarize(n = n(),
            mean = mean(salary),
            sd = sd(salary),
            se = sd / sqrt(n))
```

## Multivariate graphs|Combining grouping and faceting

```{r}
# Check the categories of discipline
levels(salaries$discipline) # A = Theoretical, B = Applied
# Create better labels for discipline's categories
sal_srd$discipline <- factor(sal_srd$discipline, levels = c("A", "B"),
                              labels = c("Theoretical","Applied"))
sal_srd
```

## Multivariate graphs|Combining grouping and faceting

```{r, eval=FALSE}
# Create plot
ggplot(sal_srd, aes(x = sex, y = mean, color = sex)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = .1) +
  scale_y_continuous(breaks = seq(70000, 140000, 10000), label = scales::dollar) +
  facet_grid(. ~ rank + discipline) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x="", y="", 
       title="Nine month academic salaries by gender, discipline, and rank",
       subtitle = "(Means and standard errors)") +
  scale_color_brewer(palette="Set1")
```

## Multivariate graphs|Combining grouping and faceting

```{r, echo=FALSE, fig.width=9, fig.height=4.3}
# Create plot
ggplot(sal_srd, aes(x = sex, y = mean, color = sex)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, 
                    ymax = mean + se),
                width = .1) +
  scale_y_continuous(breaks = seq(70000, 140000, 10000),
                     label = scales::dollar) +
  facet_grid(. ~ rank + discipline) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x="", 
       y="", 
       title="Nine month academic salaries by gender, discipline, and rank",
       subtitle = "(Means and standard errors)") +
  scale_color_brewer(palette="Set1")
```


## Multivariate graphs|Exercises (all)



# Maps

## Maps|Introduction (Joe)


## Maps|Dot density maps (Arsenio)

* Use points on a map to explore spatial relationships
* The geographic coordinates (longitude and latitude) are usually required for each observation
* E.g.: Suppose we have located moquitoes' habitats in Manhica district, and recorded their geographic coordinates and the number of anophel larvae
  + Let's plot the  locations of the habitats on the map

```{r, message=FALSE, warning=FALSE}
library(leaflet)
mosquito <- rio::import("data/mosquito_habitat.rds") # Note: This are fake data
# Print the variable names
names(mosquito)
```

## Maps|Dot density maps

```{r}
mosquito[4:6,]
```


* Plot mosquitoes' habitat on the map using circles markers

```{r, eval=FALSE}
library(leaflet)
leaflet(mosquito, width = 1000, height = 600) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng = ~longitude, lat = ~latitude)
```


***

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(leaflet)
leaflet(mosquito, width = 1000, height = 600) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng = ~longitude, lat = ~latitude)
```


## Maps|Dot density maps -- enhancements

```{r, eval=FALSE}
leaflet(mosquito, width = 1000, height = 600) %>%
  addTiles() %>%  
  addCircleMarkers(lng = ~longitude,lat = ~latitude,
                   color = ~ifelse(anophel_larvae>0,'red','green'),
                   popup = ~paste0('Habitat: ', habitat_id,
                          ', Number of anophel larvae: ', anophel_larvae),
                   label = ~habitat_id,
                   radius = 7,stroke = F,fillOpacity = 0.5)

```


***

```{r, echo=FALSE}
leaflet(mosquito, width = 1000, height = 600) %>%
  addTiles() %>%  
  addCircleMarkers(lng = ~longitude,lat = ~latitude,
                   color = ~ifelse(anophel_larvae>0,'red','green'),
                   popup = ~paste0('Habitat: ', habitat_id,
                          ', Number of anophel larvae: ', anophel_larvae),
                   label = ~habitat_id,
                   radius = 7,stroke = F,fillOpacity = 0.5)

```


## Maps|Choropleth maps (Ben)

- A choropleth map displays divided geographical areas or regions that are coloured in relation to a numeric variable. 

```{r}
data(quakes)
dat <- quakes
# Show first 20 rows from the `quakes` dataset
leaflet(data = dat) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

## Maps|Exercises (all)
