---
title: "Exercises"
author: "DataBrew"
output: 
    html_document:
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

1. Load up data on health and sociodemographic factors from Switzerland.

```{r}
data('swiss')
swiss <- swiss
```

2. Look at the `cabeça` of the data. Describe it.

3. Make a chart showing the association between Agriculture rate and Fertility rate.

4. Describe the association.

5. Create a variable called `canton`. Assign to this variable the `row.names` of the data. To do this:

```{r}
swiss$canton <- row.names(swiss)
```

6. Create an object called `x`. This will be the `swiss` data, but only keeping those cantons which are more than 80% catholic.

7. How many rows are in `x`?

8. Go back to your `swiss` data. Is there an association between `Education` and `Catholicism`? Show it in a chart.

9. What is the relationship between infant mortality and education? Show it in a chart.

10. Arrange the data by infant mortality. Which canton has the lowest?

11. What is the maximum infant mortality for all Swiss canons?

12. Make a "density" plot of fertility.

13. Create a new variable in `swiss` using `mutate`. This should be whether a canton is very agricultural or not (ie, more than 50%). To do this:

```{r}
swiss <- swiss %>%
  mutate(very_ag = Agriculture > 50)
```

14. Create an overlapping density chart of infant mortality in which the `fill` shows whether a place is very agricultural or not.

15. Create a new variable in your data called `very_cath`. This should whether or not a canton is very Catholic (ie, > 50%).


16. Create an overlapping density chart of fertility in which the `fill` is whether a place is very Catholic or not.