---
title: "Exercises 1 b"
author: "DataBrew"
output: 
    html_document:
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load up some data on insecticide by running the following:

```{r}
data('InsectSprays')
insect <- InsectSprays
```


2. Create overlapping histograms by spray

3. Create a distribution plot. Use `facet_wrap()` to create separate panels for each spray

4. Create a new variable called `avg`. This will be the average number of seconds to death for each spray.

5. Create a new variable called `slow`. This will be whether the time to death was slower than the average time to death for each spray.

6. Create a new variable called `new_spray`. This will be whether the spray is a vowel or a consonsant.

7. Create an overlapping distribution plot showing the times to death for consonant vs. vowel sprays.

8. Create a point distribution plot (use `geom_jitter`) with the `new_spray` on the x axis and the time to death on the y axis.

9. Add a `geom_violin` to the above chart.