---
title: "Exercises (Tuesday afternoon)"
author: "DataBrew"
output: 
    html_document:
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Part 1

1. Load up data on health and sociodemographic factors from Switzerland.

```{r}
data('swiss')
swiss <- swiss
```

2. Look at the `cabeça` of the data. Describe it.

3. Make a chart showing the association between Agriculture rate and Fertility rate.

4. Describe the association.

5. Create a variable called `canton`. Assign to this variable the `row.names` of the data. To do this:

```{r}
swiss$canton <- row.names(swiss)
```

6. Create an object called `x`. This will be the `swiss` data, but only keeping those cantons which are more than 80% catholic.

7. How many rows are in `x`?

8. Go back to your `swiss` data. Is there an association between `Education` and `Catholicism`? Show it in a chart.

9. What is the relationship between infant mortality and education? Show it in a chart.

10. Arrange the data by infant mortality. Which canton has the lowest?

11. What is the maximum infant mortality for all Swiss canons?

12. Make a "density" plot of fertility.

13. Create a new variable in `swiss` using `mutate`. This should be whether a canton is very agricultural or not (ie, more than 50%). To do this:

```{r}
swiss <- swiss %>%
  mutate(very_ag = Agriculture > 50)
```

14. Create an overlapping density chart of infant mortality in which the `fill` shows whether a place is very agricultural or not.

15. Create a new variable in your data called `very_cath`. This should whether or not a canton is very Catholic (ie, > 50%).

16. Create an overlapping density chart of fertility in which the `fill` is whether a place is very Catholic or not.

## Part 2

17. Load up some data about the weight of chickens over time. Create an object called `frangos`.

```{r}
data('ChickWeight')
frangos <- ChickWeight
```

18. How many columns does the data have?

19. How many rows does the data have?

20. What are the variable types (quantitative/numeric or categorical)?

21. Create a point chart showing weight on the x-axis and time on the y-axis.

22. Create an object called `frango1`. This should be just the data for chicken number 1 (ie, 1 in the `Chick` column).

23. Chart the weight of chicken 1 over time using `geom_point()`

23. Chart the weight of chicken 1 over time using `geom_line()`

24. Chart the weight of chicken 1 over time using `geom_area()`

25. Chart the weight of _all_ chickens over time using `geom_line()`. Make the color of each line different for each chicken.

26. Create an object called `zero`. This should be the frangos only at day 0 (ie, when they are born).

27. Make a histogram of chicken's weights at day 0.

28. Make a density plot of chicken's weights at day 0.

29. Make a violin plot of chicken's weights at day 0. On the x-axis, put the `Diet` type. On the y-axis, put the weight.

30. Add points to the violin plot

## Part 3

31. Run the following to get some data on "unidades sanitarias" on your computer:
```{r}
us <- read_csv("https://raw.githubusercontent.com/databrew/databrew.github.io/master/us.csv")
```

32. Look at the head of hte data.

33. Make a scatterplot of the `us` data, with `longitude` on the x-axis and `latitude` on the y-axis.

34. Make an object called `province`. This should be the number of unidades sanitarias per province.

35. Make an object called `district`. This should be the number of unidades sanitarias per district.

36. Create an object called `types`. This should be the number of unidades sanitarias per type

37. Make a barplot of the number of unidades sanitarias by type.

38. Add `coord_flip` to the code to make the plot horizontal instead of vertical.

39. Which type of health post is most common?
 
40. Make another a scatterplot of the `us` data, with `longitude` on the x-axis and `latitude` on the y-axis. But make the color of each point be a reflection of the province
