---
title: "Data Visualization using R"
subtitle: "Focus on ggplot2"
author: "Arsénio Nhacolo, Joe Brew and Ben Brew"
date: "02-06/12/2019"
output: 
  ioslides_presentation:
    logo: logo_full.png
    css: cism_styles.css
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
npac <- c("ggplot2","plotly" , "dplyr", "tidyr", "mosaicData", "carData", "rio", "VIM", 
        "scales", "treemapify", "gapminder", "ggmap", "choroplethr", "choroplethrMaps", 
        "CGPfunctions", "ggcorrplot", "visreg", "gcookbook", "forcats", "survival", 
        "survminer", "ggalluvial", "ggridges", "GGally", "superheat", "waterfalls", 
        "factoextra", "networkD3", "ggthemes", "hrbrthemes", "ggpol", "ggbeeswarm",
        "knitr", "kableExtra")
for (pk in npac){
  library(pk,character.only = TRUE)
}

```

# Outline

## Outline 

<div style="float: left; width: 50%;">
* Preparing the environment 
* Preparing the data 
* Introduction to ggplot2
* Univariate graphs
* Bivariate graphs
* Multivariate graphs
* Maps
</div>

<div style="float: right; width: 50%;">
* Time-dependent graphs
* Statistical models
* Other graphs
* Customizing graphs
* Saving graphs
* Interactive graphs
* Advice
</div>



# Preparing the environment

## Preparing the environment|Installing R and RStudio

For this training, we will use:

1. **R** -- a free software environment for statistical computing and graphics (download and install from https://cloud.r-project.org/)
2. **RStudio** -- an integrated development environment for R (download and install from https://rstudio.com/products/rstudio/download/)
3. **Packages** -- extend the capabilities of R. To install a package, open RStudio then
   + Packages --> Install --> *type the package name in Packages* --> Install or
   + Type `install.packages("package_name")` then hit Enter key
     - e.g.: install.packages("plotly")

  
## Preparing the environment|Installing the necessary packages

Run the following code to install all the necessary packages
```{r, eval=FALSE}
npac <- c("ggplot2","plotly" , "dplyr", "tidyr", "mosaicData", "carData", "rio", "VIM", 
        "scales", "treemapify", "gapminder", "ggmap", "choroplethr", "choroplethrMaps", 
        "CGPfunctions", "ggcorrplot", "visreg", "gcookbook", "forcats", "survival", 
        "survminer", "ggalluvial", "ggridges", "GGally", "superheat", "waterfalls", 
        "factoextra", "networkD3", "ggthemes", "hrbrthemes", "ggpol", "ggbeeswarm",
        "knitr", "kableExtra")

install.packages(npac)
```


# Preparing the data

## Preparing the data|Importing data from files

Different packages/functions to import data from different file formats

<div style="float: left; width: 50%;">
```{r, eval=FALSE}
library(readr)
# from a comma delimited file
opd <- read_csv("opd.csv")
# from a tab delimited file
opd <- read_tsv("opd.txt")

library(readxl)
# from an Excel workbook
opd <- read_excel("opd.xlsx", sheet=1)
```
</div>


<div style="float: right; width: 50%;">
```{r, eval=FALSE}
library(haven)
# from Stata file
opd <- read_dta("opd.dta")
# from SPSS file
opd <- read_sav("opd.sav")
# from SAS file
opd <- read_sas("opd.sas7bdat")
```
</div>

## Preparing the data|Importing data from files

Swiss-army knife for data import/export: `rio`
```{r, eval=FALSE}
library(rio)
# from any (supported) file format 
opd <- import("opd.csv") 
opd <- import("opd.txt")
opd <- import("opd.xlsx")
opd <- import("opd.dta")
opd <- import("opd.sav")
opd <- import("opd.sas7bdat")
```


## Preparing the data|Importing data from databases management systems

Different packages for different DBMS, e.g.:

* RODBC --  implements ODBC, an API for accessing DBMS
* RMySQL/RMariaDB -- for accessing MySQL/MariaDB databases
* ROracle -- for accessing Oracle database
* RPostgreSQL -- for accessing PostgreSQL database
* RSQLite -- for accessing SQLite database
* RMongo -- for accessing MongoDB database
* ...

## Preparing the data|Importing data from databases management systems

E.g.: Importing data from MariaDB

```{r, eval=FALSE}
library(RMariaDB)

con <- dbConnect(RMariaDB::MariaDB(), user = 'user_name', password  = 'password',
                 host = 'server_address', dbname = 'database_name')

death <- dbGetQuery(con,"SELECT d.person_id, m.date_birth, d.date_death, 
                    DATEDIFF(d.date_death,m.date_birth) AS age_days, 
                    d.place_death, d.place_death_other 
                    FROM dsdb.death AS d LEFT JOIN dsdb.resident AS m 
                    ON d.person_id=m.person_id 
                    WHERE YEAR(d.date_death) IN (2017, 2018) 
                    HAVING age_days BETWEEN 28 AND 1800")
```


## Preparing the data|Cleaning data

* Often the most time-consuming part of any data analysis
* Some useful functions:



```{r,echo=FALSE}
Package <- c("dplyr", "dplyr", "dplyr", "dplyr", "dplyr", "tidyr", "tidyr")
Function <- c("select", "filter", "mutate", "summarize", "group_by", "gather", "spread")
Use <- c("select variables/columns",
         "select observations/rows",
         "transform or recode variables",
         "summarize data",
         "identify subgroups for further processing",
         "convert wide format dataset to long format",
         "convert long format dataset to wide format")

tibble(Package, Function, Use) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c('striped', 'hover', 'condensed'),full_width = TRUE,
                position = 'center')
```


## Preparing the data|Tranforming data



# Introduction to ggplot2

## Introduction to ggplot2



# Univariate graphs

## Univariate graphs|Categorical

## Univariate graphs|Quantitative



# Bivariate graphs

## Bivariate graphs|Categorical vs Categorical

## Bivariate graphs|Quantitative vs Quantitative

## Bivariate graphs|Categorical vs Quantitative



# Multivariate graphs

## Multivariate graphs|Grouping



# Maps

## Maps|Dot density maps

## Maps|Choropleth maps



# Time-dependent graphs

## Time-dependent graphs|Time series

## Time-dependent graphs|Dummbbell charts

## Time-dependent graphs|Slope graphs

## Time-dependent graphs|Area charts



# Statistical models

## Statistical models|Correlation plots

## Statistical models|Linear regression

## Statistical models|Logistic regression

## Statistical models|Survival plots

## Statistical models|Mosaic plots



# Other graphs

## Other graphs|3-D scatterplot

## Other graphs|Biplots

## Other graphs|Bubble charts

## Other graphs|Flow diagrams

## Other graphs|Heatmaps

## Other graphs|Radar charts

## Other graphs|Scatterplot matrix

## Other graphs|Waterfall charts

## Other graphs|Word clouds



# Customizing graphs

## Customizing graphs|Axes

## Customizing graphs|Colours

## Customizing graphs|Points and lines

## Customizing graphs|Legends

## Customizing graphs|Labels

## Customizing graphs|Annotations

## Customizing graphs|Themes

## Customizing graphs|BBC-like (bbplot)



# Saving graphs

## Saving graphs|Via menus

## Saving graphs|Via code

## Saving graphs|File formats

## Saving graphs|External editing



# Interactive graphs

## Interactive graphs|leaflet

## Interactive graphs|plotly

## Interactive graphs|rbokeh

## Interactive graphs|rCharts

## Interactive graphs|highcharter



# Advice

## Advice|Labeling

## Advice|Signal to noise ratio

## Advice|Color choice

## Advice|y-Axis scaling

## Advice|Attribution

## Advice|Going further

## Advice|Final note



# References

## References

- Kabacoff, R. (2018). *Data Visualization with R*. QAC
- Moulik, T. (2018). *Applied Data Visualization with R and Ggplot2*. Packt Publishing









## {data-background=3.png data-background-size=cover}
